Tumor Response GAM models with comments

### 2021-05-04 Code and miscellaneous/Tumor_response_GAM__comments.Rmd
Compiled pdf for RMarkdown document of GAM model for tumor data

---
title: "Tumor_Response_GAM_comments"
author: "Ariel Mundo"
date: "5/2/2021"
output: html_document
---

```{r,loading packages}
library(mgcv)
library(tidyverse)
library(ggplot2)
library(patchwork)
library(gratia)
library (ggsci)
```

@@

```{r,simulating data}
dat <- read_csv(here::here("data", "Vishwanath StO2.csv"))
dat$Group<-as.factor(dat$Group)

@@

n <- 10 #number of observations
sd <- 10 #approximate sd from paper
set.seed(1) #setting seed
df <- 6
dat_sim <- simulate_data(dat, n, sd)

@@

#Model with interaction using s
mod_int2 <- gam(StO2_sim ~ Group+s(Day, by = Group, k = 5), data  = dat_sim)
appraise(mod_int2)
```

@@

"To avoid confounding problems the smooths are all subject to sum to zero constraints, which usually means that the main effect of g should also be included
in the model specification. For example, g + te(z,x,by=g)(pg.327)" (From Simon Wood's book)

@@

```{r, plotting model}
#plotting smooths and simulated data

@@

```{r,missing data}
    ##################################################################
############################################################
#missing data
#create a sequence of 40 random numbers between 1 and 100, these numbers will
#correspond to the row numbers to be randomly erased from the original dataset
missing <- sample(1:100, 40)

@@

c2<-ggplot(comp_StO2_m, aes(x = Day, y = diff, group = pair)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
    geom_line(color='black',size=1) +
    facet_wrap(~ pair) +
    labs(x = 'Days', y = expression(paste('Difference in StO'[2],'(missing data)' )))+
  scale_x_continuous(breaks=c(0,2,5,7,10))+
  theme_classic()+
  theme(
     text=element_text(size=18),
     legend.title=element_blank()
     )

@@

>>>>

---
title: "Tumor Response GAM models with comments"
author: "Ariel Mundo"
date: "5/2/2021"
output:
  pdf_document: default
  html_document: default
---

```{r,loading packages,echo=FALSE,include=FALSE,message=FALSE,warning=FALSE}
library(mgcv)
library(tidyverse)
library(ggplot2)
library(patchwork)
library(gratia)
library (ggsci)
```

@@

```{r,simulating data,cho=FALSE,include=FALSE,message=FALSE,warning=FALSE}
dat <- read_csv(here::here("data", "Vishwanath StO2.csv"))
dat$Group<-as.factor(dat$Group)

@@

n <- 10 #number of observations
sd <- 10 #approximate sd from paper
set.seed(1) #set seed for reproducibility
df <- 6
dat_sim <- simulate_data(dat, n, sd)

@@

#Model with interaction using s
mod_int2 <- gam(StO2_sim ~ Group+s(Day, by = Group, k = 5),method='REML', data  = dat_sim)
appraise(mod_int2)
```

@@

"To avoid confounding problems the smooths are all subject to sum to zero constraints, which usually means that the main effect of g should also be included
in the model specification. For example, g + te(z,x,by=g)(pg.327)" (From Simon Wood's book). This gives a good explanation on why mod_int1 and mod_int2 work.

@@

```{r, plotting model,echo=FALSE,include=FALSE,message=FALSE,warning=FALSE}
#plotting smooths and simulated data

@@

```{r,missing data,echo=FALSE,include=TRUE,message=FALSE,warning=FALSE}
    ##################################################################
############################################################
#missing data
#create a sequence of 40 random numbers between 1 and 100, these numbers will
#correspond to the row numbers to be randomly erased from the original dataset
missing <- sample(1:100, 40)

@@

c2<-ggplot(comp_StO2_m, aes(x = Day, y = diff, group = pair)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
    geom_line(color='black',size=1) +
    facet_wrap(~ pair) +
    labs(x = 'Days', 
         y = expression(paste('Difference in StO'[2],'\n (missing data)' 
                              )))+
  scale_x_continuous(breaks=c(0,2,5,7,10))+
  theme_classic()+
  theme(
     text=element_text(size=18),
     legend.title=element_blank()
     )

@@

So two questions remains after this:
-How do we want to present the pairwise smooth confidence interval comparison? 
The GAMs are referred as "empirical Bayesian", but I could not find a straightforward reference that had information in detail of what the pairwise comparisons can tell at certain timepoints. Sure, they are different, but what else?

-Do we want to address prediction? I recently saw a webinar from Gavin Simpson regarding this, wondered if it would be feasible to include in a concise manner.
